---
title: "Spring 2019 Term Project"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(zoo) 
library(xts)
library(plotly)
library(YieldCurve) # this is where the data comes from
FedYieldCurve <- read.csv(file.choose(), stringsAsFactors = FALSE) # this is a separate variable with the 30y yields, load from the blackboard file
data(FedYieldCurve) # load data to be used. Be sure to take a look at it
DGS30 <- read.csv(file.choose(), stringsAsFactors = FALSE) # this is a separate variable with the 30y yields, load from the blackboard file
cor(FedYieldCurve) # correlation of the rates
df <- FedYieldCurve # this dataframe has yields upt to 10y. You may want to construct a second dataframe that inclused yields to 30y
```

## Identify cycles of interest rate levels across time 

After plot 3-month, 1-year and 10 years rate across time,it shows that 3-month and 1-year have 5 interest-rate phases and 10 years has 4 interest-rate phases.


```{r plots}
plot(FedYieldCurve[,c(1,3,8)])
```


## Fit Nelson-Siegel model to each of these cycles


#1-58
```{r 1, echo=FALSE}
start1 <- 1  
end1 <- 58  # number of periods to fit
rate.Fed1 <- FedYieldCurve[start:end,]
l1 <- 3.7    # lambda is 1/tao from original paper, and t is m (maturity) 
t_ <- c(3/12, 0.5, 1,2,3,5,7,10)
t <- matrix(t_,nrow=nrow(rate.Fed),ncol=length(t_),byrow=TRUE)
slope <- (1-exp(-l*t))/(l*t) # make the curvature variable
curvature <- (1-exp(-l*t))/(l*t) - 1-exp(-l*t)
slope_single_column <- as.vector(slope) # do the same for curvature and rates. We want the data like this to fit a linear regression.
curvature_single_column <- as.vector(curvature)
rates_single_column <- as.vector(rate.Fed)
NS_OLS1 <- lm(rates_single_column ~ slope_single_column + curvature_single_column)
summary(NS_OLS1)
```

#59-145
```{r 2, echo=FALSE}
start2 <- 59  
end2 <- 145  # number of periods to fit
rate.Fed2 <- FedYieldCurve[start:end,]
l2 <- 3.8    # lambda is 1/tao from original paper, and t is m (maturity) 
t_ <- c(3/12, 0.5, 1,2,3,5,7,10)
t <- matrix(t_,nrow=nrow(rate.Fed),ncol=length(t_),byrow=TRUE)
slope <- (1-exp(-l*t))/(l*t) # make the curvature variable
curvature <- (1-exp(-l*t))/(l*t) - 1-exp(-l*t)

slope_single_column <- as.vector(slope) # do the same for curvature and rates. We want the data like this to fit a linear regression.
curvature_single_column <- as.vector(curvature)
rates_single_column <- as.vector(rate.Fed)

NS_OLS2 <- lm(rates_single_column ~ slope_single_column + curvature_single_column)
summary(NS_OLS)
```

#146-269
```{r 3, echo=FALSE}
start3 <- 146  
end3 <- 269  # number of periods to fit
rate.Fed3 <- FedYieldCurve[start:end,]
l3 <- 3.8    # lambda is 1/tao from original paper, and t is m (maturity) 
t_ <- c(3/12, 0.5, 1,2,3,5,7,10)
t <- matrix(t_,nrow=nrow(rate.Fed),ncol=length(t_),byrow=TRUE)
slope <- (1-exp(-l*t))/(l*t) # make the curvature variable
curvature <- (1-exp(-l*t))/(l*t) - 1-exp(-l*t)

slope_single_column <- as.vector(slope) # do the same for curvature and rates. We want the data like this to fit a linear regression.
curvature_single_column <- as.vector(curvature)
rates_single_column <- as.vector(rate.Fed)

NS_OLS3 <- lm(rates_single_column ~ slope_single_column + curvature_single_column)
summary(NS_OLS3)
```

#270-324
```{r 4, echo=FALSE}
start4 <- 270  
end4 <- 324  # number of periods to fit
rate.Fed4 <- FedYieldCurve[start:end,]
l4 <- 3.9   # lambda is 1/tao from original paper, and t is m (maturity) 
t_ <- c(3/12, 0.5, 1,2,3,5,7,10)
t <- matrix(t_,nrow=nrow(rate.Fed),ncol=length(t_),byrow=TRUE)
slope <- (1-exp(-l*t))/(l*t) # make the curvature variable
curvature <- (1-exp(-l*t))/(l*t) - 1-exp(-l*t)

slope_single_column <- as.vector(slope) # do the same for curvature and rates. We want the data like this to fit a linear regression.
curvature_single_column <- as.vector(curvature)
rates_single_column <- as.vector(rate.Fed)

NS_OLS4 <- lm(rates_single_column ~ slope_single_column + curvature_single_column)
summary(NS_OLS)
```

I used Adjusted R-squared to find the best lambda for each curve.
##plot the real, observed yield curve alongside fitted curve for each cycle


```{r 11, echo=FALSE}
nelson_siegel_calculate<-function(lambda,tau,beta0,beta1,beta2){
  beta0 + beta1*(1-exp(-lambda*tau))/(lambda*tau) + beta2*((1-exp(-lambda*tau))/(lambda*tau) - exp(-lambda*tau))
}

# Plot the yield curve of NS with the following parameters
maturity <- c(.25,.5,1,2,3,5,7,10)
tau <- maturity
lambda <-l1  # this is really tau=200 from NS paper, since rates are multiplied by 100
beta0 <- as.numeric(NS_OLS1$coef[1])
beta1 <- as.numeric(NS_OLS1$coef[2])
beta2 <- as.numeric(NS_OLS1$coef[3])

ns_data <- data.frame(maturity, ns_yield=nelson_siegel_calculate(tau,lambda,beta0,beta1,beta2))
head(ns_data) #forecast yield curve

# Plot observed and fitted curve
rates <- FedYieldCurve[end1,] # Depends on the period to be analyzed, this is just one period
rates<-as.vector(rates)
df <- data.frame(maturity,rates) #observed yield curve
ggplot(df, aes(x=maturity, y=rates)) + geom_point() + geom_line(data=ns_data)

ggplot() + 
  geom_line(data=df, aes(x=maturity, y=rates, colour="Observed")) + 
  geom_line(data=ns_data, aes(x=maturity, y=ns_yield, colour="Fitted")) + labs(title="Observed vs Fitted Curve")

# Goodness of fit of the model (MSE)
mse <- (1/(end1-start1))*sum((df[,2] - ns_data[,2]) ^ 2)
cat("With Lambda =",lambda,"MSE between two curves is:",mse)
# what about r-squared?
tss <- sum((ns_data[,2] - mean(ns_data[,2])) ^ 2)
rss <- sum((ns_data[,2] - df[,2]) ^ 2)
rsq <- 1 - rss/tss # calculate rss and tss
cat("R-squared is: ",rsq)
```



```{r 22, echo=FALSE}
nelson_siegel_calculate<-function(lambda,tau,beta0,beta1,beta2){
  beta0 + beta1*(1-exp(-lambda*tau))/(lambda*tau) + beta2*((1-exp(-lambda*tau))/(lambda*tau) - exp(-lambda*tau))
}

# Plot the yield curve of NS with the following parameters
maturity <- c(.25,.5,1,2,3,5,7,10)
tau <- maturity
lambda <-l2  # this is really tau=200 from NS paper, since rates are multiplied by 100
beta0 <- as.numeric(NS_OLS2$coef[1])
beta1 <- as.numeric(NS_OLS2$coef[2])
beta2 <- as.numeric(NS_OLS2$coef[3])

ns_data <- data.frame(maturity, ns_yield=nelson_siegel_calculate(tau,lambda,beta0,beta1,beta2))
head(ns_data) #forecast yield curve

# Plot observed and fitted curve
rates <- FedYieldCurve[end2,] # Depends on the period to be analyzed, this is just one period
rates<-as.vector(rates)
df <- data.frame(maturity,rates) #observed yield curve
ggplot(df, aes(x=maturity, y=rates)) + geom_point() + geom_line(data=ns_data)

ggplot() + 
  geom_line(data=df, aes(x=maturity, y=rates, colour="Observed")) + 
  geom_line(data=ns_data, aes(x=maturity, y=ns_yield, colour="Fitted")) + labs(title="Observed vs Fitted Curve")

# Goodness of fit of the model (MSE)
mse <- (1/(end2-start2))*sum((df[,2] - ns_data[,2]) ^ 2)
cat("With Lambda =",lambda,"MSE between two curves is:",mse)
# what about r-squared?
tss <- sum((ns_data[,2] - mean(ns_data[,2])) ^ 2)
rss <- sum((ns_data[,2] - df[,2]) ^ 2)
rsq <- 1 - rss/tss # calculate rss and tss
cat("R-squared is: ",rsq)
```

```{r 33, echo=FALSE}
nelson_siegel_calculate<-function(lambda,tau,beta0,beta1,beta2){
  beta0 + beta1*(1-exp(-lambda*tau))/(lambda*tau) + beta2*((1-exp(-lambda*tau))/(lambda*tau) - exp(-lambda*tau))
}

# Plot the yield curve of NS with the following parameters
maturity <- c(.25,.5,1,2,3,5,7,10)
tau <- maturity
lambda <-l3  # this is really tau=200 from NS paper, since rates are multiplied by 100
beta0 <- as.numeric(NS_OLS3$coef[1])
beta1 <- as.numeric(NS_OLS3$coef[2])
beta2 <- as.numeric(NS_OLS3$coef[3])

ns_data <- data.frame(maturity, ns_yield=nelson_siegel_calculate(tau,lambda,beta0,beta1,beta2))
head(ns_data) #forecast yield curve

# Plot observed and fitted curve
rates <- FedYieldCurve[end3,] # Depends on the period to be analyzed, this is just one period
rates<-as.vector(rates)
df <- data.frame(maturity,rates) #observed yield curve
ggplot(df, aes(x=maturity, y=rates)) + geom_point() + geom_line(data=ns_data)

ggplot() + 
  geom_line(data=df, aes(x=maturity, y=rates, colour="Observed")) + 
  geom_line(data=ns_data, aes(x=maturity, y=ns_yield, colour="Fitted")) + labs(title="Observed vs Fitted Curve")

# Goodness of fit of the model (MSE)
mse <- (1/(end3-start3))*sum((df[,2] - ns_data[,2]) ^ 2)
cat("With Lambda =",lambda,"MSE between two curves is:",mse)
# what about r-squared?
tss <- sum((ns_data[,2] - mean(ns_data[,2])) ^ 2)
rss <- sum((ns_data[,2] - df[,2]) ^ 2)
rsq <- 1 - rss/tss # calculate rss and tss
cat("R-squared is: ",rsq)
```

```{r 44, echo=FALSE}
nelson_siegel_calculate<-function(lambda,tau,beta0,beta1,beta2){
  beta0 + beta1*(1-exp(-lambda*tau))/(lambda*tau) + beta2*((1-exp(-lambda*tau))/(lambda*tau) - exp(-lambda*tau))
}

# Plot the yield curve of NS with the following parameters
maturity <- c(.25,.5,1,2,3,5,7,10)
tau <- maturity
lambda <-l4  # this is really tau=200 from NS paper, since rates are multiplied by 100
beta0 <- as.numeric(NS_OLS4$coef[1])
beta1 <- as.numeric(NS_OLS4$coef[2])
beta2 <- as.numeric(NS_OLS4$coef[3])

ns_data <- data.frame(maturity, ns_yield=nelson_siegel_calculate(tau,lambda,beta0,beta1,beta2))
head(ns_data) #forecast yield curve

# Plot observed and fitted curve
rates <- FedYieldCurve[end4,] # Depends on the period to be analyzed, this is just one period
rates<-as.vector(rates)
df <- data.frame(maturity,rates) #observed yield curve
ggplot(df, aes(x=maturity, y=rates)) + geom_point() + geom_line(data=ns_data)

ggplot() + 
  geom_line(data=df, aes(x=maturity, y=rates, colour="Observed")) + 
  geom_line(data=ns_data, aes(x=maturity, y=ns_yield, colour="Fitted")) + labs(title="Observed vs Fitted Curve")

# Goodness of fit of the model (MSE)
mse <- (1/(end4-start4))*sum((df[,2] - ns_data[,2]) ^ 2)
cat("With Lambda =",lambda,"MSE between two curves is:",mse)
# what about r-squared?
tss <- sum((ns_data[,2] - mean(ns_data[,2])) ^ 2)
rss <- sum((ns_data[,2] - df[,2]) ^ 2)
rsq <- 1 - rss/tss # calculate rss and tss
cat("R-squared is: ",rsq)
```